Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area.

### 2022-08-04 outputs/paper/paper.qmd
Updates to paper and presentation

---
#### Preamble ####
title: "Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area."
author: "Sarah Mansoor"
email: "sarah.mansoor@mail.utoronto.ca"
thanks: "Code and data are available at: https://github.com/sarahmansoorr/SUDS2022"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: pdf
editor: visual
abstract: "The growing impact of climate change in Canada demonstrates a clear and current threat to human health. These threats are prevalent in Canada as the country continues to face the consequences of warming almost at twice the global rate (Canada Warming Twice as Fast as the Rest of the World, BBC). The focus for this paper is on analyzing trends in temperature increases, pollution levels and the health effects of these in Toronto for the past 20 years. Impactful change can be made with this research in hopes of reversing some effects of climate change. Once the data for this research was collected, the data was analyzed using exploratory analysis, data visualization, and regression analysis. After analysis using these methods, trends and patterns show that temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO (Nitric Oxide), NO2 (Nitrogen Dioxide), CO (Carbon Monoxide), and SO2 (Sulfur Dioxide) have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months which may correlate with the use of more electricity and heating during colder months. O3 (Ozone) pollution levels have remained mostly the same since 2002 and these levels are higher in summer months which may be due to the hotter temperatures in warmer months. Temperature seems to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels. The presence of these pollutants has effects on health for the Toronto population including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function."
toc: true
bibliography: references.bib
---

@@

Urgent action is needed and the science shows that it will make a tangible difference [@]. The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada [@]. Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change [@].

>>>>

---
#### Preamble ####
title: "Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area."
author: "Sarah Mansoor"
email: "sarah.mansoor@mail.utoronto.ca"
thanks: "Code and data are available at: https://github.com/sarahmansoorr/SUDS2022"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: pdf
editor: visual
abstract: "Climate change has a growing impact in Canada. This paper analyses the trends in temperature and pollution levels in Toronto over the last 20 years using exploratory analysis, data visualization, and regression analysis. Temperatures have continued to increase since 2002 and winters are experiencing less extreme cold while summers are experiencing more extreme heat. Further, pollution levels for NO (Nitric Oxide), NO2 (Nitrogen Dioxide), CO (Carbon Monoxide), and SO2 (Sulfur Dioxide) have decreased since 2002 which may be related to laws and regulations. Data visualizations showed that these pollutants appear to have higher levels in winter months while O3 (Ozone) pollution levels have remained mostly the same since 2002 and O3 levels have higher maximums in summer months. The presence of these pollutants has detrimental health effects for the Toronto population including lung irritation, increase in allergens for those with asthma, respiratory illnesses, heart disease, cardiovascular diseases, and reduced lung function. Understanding which pollutants are most problematic during particular months can enable further research which would reduce their effects on health."
toc: true
bibliography: references.bib
---

@@

Urgent action is needed and the science shows that it will make a tangible difference \[\@\]. The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada \[\@\]. Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change \[\@\].
